\texttt{Taxonomic supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-06-18 doc/incertae_sedis.tex
Typos and clean up discussion.

In addition, our attempt to define reduced exclude sets for taxonomy splits yields
a semantics that is not the only concievable semantics for \emph{incertae sedis}
taxa (Figure \ref{fig:Placing-one-incertae}).  It allows \emph{incertae sedis} siblings to inter-digitate
(Figure \ref{fig:interdigitation}), which might be a surprising result to the taxonomist.  Furthermore,
in what might seem a lack of consistency, it does \emph{not} allow \emph{incertae sedis}
taxa to intrude into each other if they become siblings through the result of
placement, as is the case for taxa W and V in Figure \ref{fig:Placing-one-incertae}c.
While each of these problems can be solved separately by adjusting the definition of the
reduces exclude sets, no solution seems to be best in all respects.  This may be a result
of the fact that split-based approaches to supertrees seek a simultaneous solution to a set
of constraints.  This simultaneity can be contrasted with a solution that is phrased
as a set of sequential operations to the taxonomy, as would be the case when (for example)
(i) placing V within WX so that it is sister to W, (ii) marking V as no longer \emph{incertae sedis}, and then (iii)
placing W within V (which is no longer \emph{incertae sedis}).  This maybe considered a ``conceptual difficulty''.

Another issue that we have postponed for future work is how to more narrowly delimit
the ranges in which \emph{incertae sedis} may attach.  The current paper allows
\emph{incertae sedis} to attach anywhere within their sisters, but the taxonomist
may only consider it plausible for the taxon to intrude only within specific
siblings.  Furthermore, our method also assumes that an \emph{incertae sedis} label on a taxon
indicates that a taxon may attach at any depth within its siblings.  In practice, an
\emph{incertae sedis} genus with sister families (for example) is unlikely to attach
so deeply within a sister family that it becomes nested within another genus.

As we demonstrated above (in the discussion of figure \ref{fig:Placing-one-incertae-MRP}),
if the \emph{incertae sedis} status of a taxon consists solely of
    a flag on that taxon, we can encounter 
    multiple equally plausible ways of encoding the information
    into a character matrix.
Resolving such ambiguities would require some datastore of intended
    meanings of each restricted \emph{incertae sedis} annotation.
In order to consider more precise attachment ranges for \emph{incertae sedis} taxa,
    we would have to obtain machine readable specifications for attachment ranges
    for groupings that a taxonomist has labeled as \emph{incertae sedis}
Obtaining these specifications would be difficult or impossible in many cases.
However,  not all \emph{incertae sedis} taxa in our taxonomy are directly
labeled \emph{incertae sedis} by taxonomists.
\emph{Incertae sedis}
taxa can also result from automatic merging of taxonomies to create
the OpenTree taxonomy.
For example, in Figure 3 of
\citet{rees2017automated}, cases \#4 and \#6 illustrate examples where
merging of two taxonomies leads to a taxonomy with a taxon of
uncertain placement.

Taxonomy merging generates a automated \emph{incertae sedis} annotation
    in a number of ways.
Perhaps, the easiest cause to contemplate is if
one input taxonomy,
$\taxonomy_{1}$, contains more levels of hierarchy than a second
input taxonomy,
$\taxonomy_{2}$.
When the merger adds must add internal nodes that are unique to $\taxonomy_{2}$,
it is unclear where these
extra leaves should be nested be placed in the richer hierarchy of $\taxonomy_1$.
In such a case, it is feasible to imagine the taxonomy merging procedure outputting
    a list of plausible attachment points for each generated \emph{incertae sedis}
    label.
Case in which the \emph{incertae sedis} label is present in an input taxonomy,
    restricting the range of attachment points would need to be treated as
    some manually curated statements.

>>>>

In addition, our attempt to define reduced exclude sets for taxonomy splits yields
a semantics that is not the only concievable semantics for \emph{incertae sedis}
taxa (Figure \ref{fig:Placing-one-incertae}).  It allows \emph{incertae sedis} siblings to inter-digitate
(Figure \ref{fig:interdigitation}), which might be a surprising result to the taxonomist.  Furthermore,
in what might seem a lack of consistency, it does \emph{not} allow \emph{incertae sedis}
taxa to intrude into each other if they become siblings through the result of
placement, as is the case for taxa W and V in Figure \ref{fig:Placing-one-incertae}c.
While each of these problems can be solved separately by adjusting the definition of the
reduced exclude sets, no solution seems to be best in all respects.  This may be a result
of the fact that split-based approaches to supertrees seek a simultaneous solution to a set
of constraints.  This simultaneity can be contrasted with a solution that is phrased
as a set of sequential operations to the taxonomy, as would be the case when (for example)
(i) placing V within WX so that it is sister to W, (ii) marking V as no longer \emph{incertae sedis}, and then (iii)
placing W within V (which is no longer \emph{incertae sedis}).  This maybe considered a ``conceptual difficulty''.

Another issue that we have postponed for future work is how to more narrowly delimit
the ranges in which \emph{incertae sedis} may attach.  The current paper allows
\emph{incertae sedis} to attach within any of their siblings, but the taxonomist
may only consider it plausible for the taxon to intrude only within specific
siblings.  We also assume that a taxon labeled as \emph{incertae sedis} may
attach at any depth within its siblings. In practice, an \emph{incertae sedis}
genus with sister families (for example) is unlikely to attach so deeply within
a sister family that it becomes nested within another genus.

Our method also assumes that the \emph{incertae sedis} status of a taxon
consists solely of a flag on that taxon.  In order to consider more precise
attachment ranges for \emph{incertae sedis} taxa, we would have to obtain
machine readable specifications for attachment ranges for groupings that a
taxonomist has labeled as \emph{incertae sedis}.  Obtaining these specifications
would be difficult or impossible in many cases. However, not all \emph{incertae
sedis} taxa in our taxonomy are directly labeled \emph{incertae sedis} by
taxonomists. \emph{Incertae sedis} taxa can also result from automatic merging
of taxonomies to create the OpenTree taxonomy. For example, in Figure 3 of
\citet{rees2017automated}, cases \#4 and \#6 illustrate examples where merging
of two taxonomies leads to a taxonomy with a taxon of uncertain placement.

Taxonomy merging generates an automated \emph{incertae sedis} annotation
in a number of ways. Perhaps, the easiest cause to contemplate is if
one input taxonomy, $\taxonomy_{1}$, contains more levels of hierarchy than a second
input taxonomy, $\taxonomy_{2}$.
When the merger adds internal nodes that are unique to $\taxonomy_{2}$,
it is unclear how these
extra leaves should be nested or placed in the richer hierarchy of $\taxonomy_1$.
In such a case, it is feasible to imagine the taxonomy-merging procedure outputting
    a list of plausible attachment points for each generated \emph{incertae sedis}
    label.
For the case in which the \emph{incertae sedis} label is present in an input taxonomy,
    restricting the range of attachment points would need to be treated as
    manually curated statements.
